Estimating Sleep Quality Metrics Using Free-Living Accelerometer Data From Thigh-Worn Devices in Comparison to an EEG-Based Sleep Tracking Device

### 2023-06-29 manuscript/manuscript.qmd
fix metrics and calculate n awakenings

In this study, raw data processing began with a low-pass filtration step
using a 4th order Butterworth filter with a 5 Hz cut-off frequency to
eliminate high-frequency noise. Following filtration, data were
partitioned into overlapping 2-second intervals, each successive
interval sharing a 50% overlap with the previous one similar to methods
described by Skotte et al.[@skotte_detection_2014]. Any non-wear data
was remove using previously described methods[@skovgaard2023] and data
was resampled to 30-second epochs so every sample classified by the
algorithms corresponds to a 30-second epoch scored during the ZM
recordings. Subsequently, we performed a feature extraction process that
yielded a set of 88 features, providing a robust characterization of the
data. Extracted from accelerometer and temperature signals, these
features include temporal elements that use both lag and lead values,
capturing dynamic data trends by incorporating measurements from
preceding and upcoming intervals. Furthermore, inspired by Walch et
al.[@walch2019], we developed sensor-independent features to encapsulate
circadian rhythms. These features offer unique insights not directly
discernible from sensor outputs and are meant to approximate the
changing drive of the circadian clock to sleep over the course of the
night (see @fig-sensor-independent). Furthermore, the feature set was
enriched by including signal characteristics, which encompass vector
magnitude, mean crossing rate, skewness, and kurtosis for each of the x,
y, and z dimensions. These features provide a comprehensive
understanding of the signal's attributes, facilitating a more nuanced
analysis of the underlying patterns. All features are summarized in
table ???. (overvej tabel til supp. material)

@@

### Model Validation \[evt bare "statistics"\]

In our study, we employed standard evaluation metrics to assess the
epoch-to-epoch performance of each model. These metrics include accuracy
as $accuracy = \frac{TP+TN}{TP+TN+FP+FN}$, sensitivity as
$sensitivity = \frac{TP}{TP+FN}$, specificity as
$specificity = \frac{TN}{TN+FP}$, precision as
$precision = \frac{TP}{TP+FP}$, F1 score as
$F_1 = 2 * \frac{precision * sensitivity}{precision + sensitivity}$, and
area under the precision-recall curves were also calculated. The Average
Precision, represented by the area under the precision-recall curve,
provides a quantitative measure of performance. A value of 1 for Average
Precision indicates the best performance. For the multiclass biLSTM
model, we calculated performance metrics as macro-averages, to treat all
classes equally. Consistent with previous studies in the
field[@hjorth2012; @kushida2001], we treated the in-bed/out-of-bed and
awake/asleep scoring tasks as a binary classification problems, where
in-bed and asleep was considered the positive label and out-of-bed and
awake the negative label.

@@

The children children icluded in the analyses were 9.4 years old (SD = 2.1).

As reported in @tbl-summary_zm the sleep quality metrics derived from ZMachine predictions were modified by the implementation of 5-minute and 10-minute median filters. The Sleep Period Time (SPT) remained consistent across raw and filtered datasets (mean: 9.2 ± 2.1 hours). Total Sleep Time (TST) and Sleep Efficiency (SE) increased in the filtered data, implying the filters categorize some wakefulness as sleep. Specifically, TST increased from a raw mean of 7.7 ± 1.9 hours to 8.1 ± 2.0 hours (5-minute filter) and 8.2 ± 2.1 hours (10-minute filter), while SE rose from 82.6 ± 12.0% to 86.4 ± 12.7% and 87.5 ± 12.9% respectively. Latency to Persistent Sleep (LPS) also elevated, suggesting the filter smooths out brief awakenings at sleep onset, leading to a prolonged time to persistent sleep. The most significant change was seen in Wake After Sleep Onset (WASO), which dropped from 39.0 ± 33.6 minutes in raw data to 30.6 ± 46.8 minutes and 22.3 ± 55.4 minutes in the 5-minute and 10-minute filtered data, respectively. The results underscore the influential role of median filters on sleep metrics.

|||Raw| |Median 5-Minute| |Median 10-Minute| |
|-|-|-|-|-|-|-|-|
|||Mean|SD|Mean|SD|Mean|SD|
||SPT|9.2|2.1|9.2|2.1|9.2|2.1|
||TST|7.7|1.9|8.1|2.0|8.2|2.1|
||SE|82.6|12.0|86.4|12.7|87.5|12.9|
||LPS|34.5|27.9|36.3|39.8|38.0|48.7|
||WASO|39.0|33.6|30.6|46.8|22.3|55.4|

: Summary of sleep quality metrics as calculated based on the raw, the 5-minute, and 10-minute median filtered raw ZM prediction. {#tbl-summary_zm}

>>>>

In this study, raw data processing began with a low-pass filtration step
using a 4th order Butterworth filter with a 5 Hz cut-off frequency to
eliminate high-frequency noise. Following filtration, data were
partitioned into overlapping 2-second intervals, each successive
interval sharing a 50% overlap with the previous one similar to methods
described by Skotte et al.[@skotte_detection_2014]. Any non-wear data
was remove using previously described methods[@skovgaard2023] and data
was resampled to 30-second epochs so every sample classified by the
algorithms corresponds to a 30-second epoch scored during the ZM
recordings. Subsequently, we performed a feature extraction process that
yielded a set of 88 features, providing a robust characterization of the
data. Extracted from accelerometer and temperature signals, these
features include temporal elements that use both lag and lead values,
capturing dynamic data trends by incorporating measurements from
preceding and upcoming intervals. Furthermore, inspired by Walch et
al.[@walch2019], we developed sensor-independent features to encapsulate
circadian rhythms. These features offer unique insights not directly
discernible from sensor outputs and are meant to approximate the
changing drive of the circadian clock to sleep over the course of the
night (see @fig-sensor-independent). Furthermore, the feature set was
enriched by including signal characteristics, which encompass vector
magnitude, mean crossing rate, skewness, and kurtosis for each of the x,
y, and z dimensions. These features provide a comprehensive
understanding of the signal's attributes, facilitating a more nuanced
analysis of the underlying patterns. All features are summarized in
table ??? (overvej tabel til supp. material)

@@

### Model Validation \[evt. bare "statistics"\]

In our study, we utilized standard evaluation metrics to assess the
performance of each model on an epoch-to-epoch basis. These include
accuracy ($accuracy = \frac{TP+TN}{TP+TN+FP+FN}$), sensitivity
($sensitivity = \frac{TP}{TP+FN}$), specificity
($specificity = \frac{TN}{TN+FP}$), precision
($precision = \frac{TP}{TP+FP}$), negative predictive value (NPV,
$NPV = \frac{TN}{TN + FN}$), and F1 score
($F_1 = 2 * \frac{precision * sensitivity}{precision + sensitivity}$).

In the context of our sequential learning strategy, the initial models
were tasked with the binary classification of in-bed vs. out-of-bed. For
this task, we assessed performance using the F1-score, accuracy,
sensitivity, specificity, and precision metrics. The second models in
our sequential learning strategy focused on the binary classification of
asleep vs. awake. For these models, we considered the same metrics, in
addition to the negative predictive rate. The class imbalance in this
case led us to compute the F1 score as an unweighted macro-average.
Additionally, we evaluated the multiclass classifier, biLSTM, using
macro-averaged F1-score, sensitivity, and precision. Furthermore, we
present precision-recall curves and receiver operating characteristic
curves, including the area under the curve, for each model and for each
type of ZM prediction.We considered both the in-bed/out-of-bed and
awake/asleep scoring tasks as binary classification problems,
designating in-bed and asleep as the positive labels and out-of-bed and
awake as the negative labels in accordance with previous
research[@hjorth2012; @kushida2001].

@@

The children children included in the analyses were 9.4 years old (SD =
2.1). The number of sleep epochs recorded in the study was 2,035,261,
which accounted for approximately 86% of the total sleep duration. This
metric indicates the granularity of sleep measurement, capturing the
discrete moments of sleep throughout the sleep period. Sleep epochs are
essential for analyzing sleep patterns and stages, as well as for
assessing sleep quality and efficiency.

As reported in @tbl-zm_overview the sleep quality metrics derived from
ZM predictions were modified by the implementation of 5-minute and
10-minute median filters. The Sleep Period Time (SPT) were consistent
across raw and filtered datasets (mean: 9.2 ± 2.1 hours) as this
corresponds to the length of the ZM recording. Total Sleep Time (TST)
and Sleep Efficiency (SE) increased in the filtered data, implying the
filters categorize some wakefulness as sleep. Specifically, TST
increased from a raw mean of 7.7 ± 1.9 hours to 8.1 ± 2.0 hours
(5-minute filter) and 8.2 ± 2.1 hours (10-minute filter), while SE rose
from 82.6 ± 12.0% to 86.4 ± 12.7% and 87.5 ± 12.9% respectively. Latency
to Persistent Sleep (LPS) also elevated, suggesting the filter smooths
out brief awakenings at sleep onset, leading to a prolonged time to
persistent sleep. The most significant change was seen in Wake After
Sleep Onset (WASO), which dropped from 39.0 ± 33.6 minutes in raw data
to 30.6 ± 46.8 minutes and 22.3 ± 55.4 minutes in the 5-minute and
10-minute filtered data, respectively. The results underscore the role
of median filters on sleep metrics.

```{r}
#| echo: false
#| message: false
#| label: tbl-zm_overview
#| tbl-cap: Overview of characteristics of the ZM sleep quality summaries. Values are represented as mean (SD).

source("code/table_overview_ZM_stats.R")

tbl_overview
```

## Performance on Epoch-To-Epoch Basis

The epoch-to-epoch evaluation of in-bed performance metrics, outlined in
@tbl-in_bed_performance, demonstrates practically equivalent performance
across all model types. While the Decision Tree model posted an F1 score
of 94.4% and accuracy of 95.3%, the Logistic Regression and Feed-Forward
Neural Network models each exhibited an F1 score of 95.0% and similar
accuracies of 95.7% and 95.8%, respectively. The XGBoost model, despite
recording the highest metrics with an F1 score of 95.4% and accuracy of
96.1%, outpaced the others only marginally. This underscores the
consistency of performance among these models in classifying in-bed
conditions.

```{r}
#| echo: false
#| message: false
#| label: tbl-in_bed_performance
#| tbl-cap: In-Bed Performance Metrics

source("code/table_in_bed_performance.R")

tbl_in_bed
```

@tbl-sleep_performance details the performance of the sequential models
on raw and median-filtered (5 and 10 minute) ZM predictions. The F1
Scores, which are unweighted macro averages, for raw ZM predictions
range from 71.05% to 76.18%. The models perform comparably, but the low
Specificity values (62.84% to 70.93%) suggest difficulty in correctly
classifying negative instances. Applying 5-minute median filtering
improves the performance metrics. The XGBoost model tops the charts with
an F1 Score of 79.22% and NPV of 74.00%. However, Specificity still
remains low, with values between 54.68% and 74.84% across all models.
With 10-minute median filtering, the metrics improve further. The
XGBoost model still leads with an F1 Score of 80.87% and an NPV of
75.76%. But, Specificity remains a concern, ranging from 57.47% to
76.35% across all models.

```{r}
#| echo: false
#| message: false
#| tbl-colwidths: auto
#| label: tbl-sleep_performance
#| tbl-cap: Sleep Performance Metrics

source("code/table_sleep_performance.R")

tbl_sleep_performance
```

The analysis of precision-recall and ROC curves across different models
and ZM prediction types shows varying performance. In terms of
precision-recall AUC, the Decision Tree model consistently outperforms
others, indicating its superior predictive accuracy (see
@fig-pr_curves). Conversely, the Neural Network model generally shows
weaker performance. However, for ROC AUC, the XGBoost model consistently
excels across all data types, indicating a strong ability to
differentiate between classes, while the Neural Network model tends to
underperform (see @fig-roc_curves). The F-measure (F1 score) shows
variable performance across different configurations but generally, the
Decision Tree model yields higher scores.

![Precision-recall curves of the models evaluated across the different
ZM predictions, including raw ZM predictions, as well as 5-minute and
10-minute median smoothing of the ZM raw predictions. The x-axis of the
plot represents the proportion of true wake epochs that were correctly
classified as wake, while the y-axis represents the proportion of all
epochs labeled as wake by the classifier that were classified correctly.
The area under the curve values are displayed as color-coded text in the
plot indicate the area under the Precision-Recall curve for each model
and condition.](visuals/plot_sleep_pr.pdf){#fig-pr_curves}

![Receiver operating characteristic curves of the models evaluated
across the different ZM predictions, including raw ZM predictions, as
well as 5-minute and 10-minute median smoothing of the ZM raw
predictions. The x-axis of the plot represents the proportion of true
asleep epochs that were incorrectly classified as awake, while the
y-axis represents the proportion of all epochs labeled as awake by the
classifier that were correctly classified. The area under the curve
values displayed are displayed as color-coded text in the plot to
indicate the area under the receiver operating characteristic curve for
each model and condition.](visuals/plot_sleep_roc.pdf){#fig-roc_curves}

The performance metrics of the biLSTM were...

### 2023-06-29 manuscript/manuscript.tex
fix metrics and calculate n awakenings

\begin{keyword}
    Sleep \sep Accelerometry \sep EEG \sep Machine learning \sep 
    Sleep quality metrics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, sharp corners, frame hidden, breakable, enhanced]}{\end{tcolorbox}}\fi

@@

In this study, raw data processing began with a low-pass filtration step
using a 4th order Butterworth filter with a 5 Hz cut-off frequency to
eliminate high-frequency noise. Following filtration, data were
partitioned into overlapping 2-second intervals, each successive
interval sharing a 50\% overlap with the previous one similar to methods
described by Skotte et al.\citep{skotte_detection_2014}. Any non-wear
data was remove using previously described methods\citep{skovgaard2023}
and data was resampled to 30-second epochs so every sample classified by
the algorithms corresponds to a 30-second epoch scored during the ZM
recordings. Subsequently, we performed a feature extraction process that
yielded a set of 88 features, providing a robust characterization of the
data. Extracted from accelerometer and temperature signals, these
features include temporal elements that use both lag and lead values,
capturing dynamic data trends by incorporating measurements from
preceding and upcoming intervals. Furthermore, inspired by Walch et
al.\citep{walch2019}, we developed sensor-independent features to
encapsulate circadian rhythms. These features offer unique insights not
directly discernible from sensor outputs and are meant to approximate
the changing drive of the circadian clock to sleep over the course of
the night (see Figure~\ref{fig-sensor-independent}). Furthermore, the
feature set was enriched by including signal characteristics, which
encompass vector magnitude, mean crossing rate, skewness, and kurtosis
for each of the x, y, and z dimensions. These features provide a
comprehensive understanding of the signal's attributes, facilitating a
more nuanced analysis of the underlying patterns. All features are
summarized in table ???. (overvej tabel til supp. material)

@@

\hypertarget{model-validation-evt-bare-statistics}{%
\subsubsection{Model Validation {[}evt bare
``statistics''{]}}\label{model-validation-evt-bare-statistics}}

In our study, we employed standard evaluation metrics to assess the
epoch-to-epoch performance of each model. These metrics include accuracy
as \(accuracy = \frac{TP+TN}{TP+TN+FP+FN}\), sensitivity as
\(sensitivity = \frac{TP}{TP+FN}\), specificity as
\(specificity = \frac{TN}{TN+FP}\), precision as
\(precision = \frac{TP}{TP+FP}\), F1 score as
\(F_1 = 2 * \frac{precision * sensitivity}{precision + sensitivity}\),
and area under the precision-recall curves were also calculated. The
Average Precision, represented by the area under the precision-recall
curve, provides a quantitative measure of performance. A value of 1 for
Average Precision indicates the best performance. For the multiclass
biLSTM model, we calculated performance metrics as macro-averages, to
treat all classes equally. Consistent with previous studies in the
field\citep{hjorth2012, kushida2001}, we treated the in-bed/out-of-bed
and awake/asleep scoring tasks as a binary classification problems,
where in-bed and asleep was considered the positive label and out-of-bed
and awake the negative label.

@@

The children children icluded in the analyses were 9.4 years old (SD =
2.1).

As reported in Table~\ref{tbl-summary_zm} the sleep quality metrics
derived from ZMachine predictions were modified by the implementation of
5-minute and 10-minute median filters. The Sleep Period Time (SPT)
remained consistent across raw and filtered datasets (mean: 9.2 ± 2.1
hours). Total Sleep Time (TST) and Sleep Efficiency (SE) increased in
the filtered data, implying the filters categorize some wakefulness as
sleep. Specifically, TST increased from a raw mean of 7.7 ± 1.9 hours to
8.1 ± 2.0 hours (5-minute filter) and 8.2 ± 2.1 hours (10-minute
filter), while SE rose from 82.6 ± 12.0\% to 86.4 ± 12.7\% and 87.5 ±
12.9\% respectively. Latency to Persistent Sleep (LPS) also elevated,
suggesting the filter smooths out brief awakenings at sleep onset,
leading to a prolonged time to persistent sleep. The most significant
change was seen in Wake After Sleep Onset (WASO), which dropped from
39.0 ± 33.6 minutes in raw data to 30.6 ± 46.8 minutes and 22.3 ± 55.4
minutes in the 5-minute and 10-minute filtered data, respectively. The
results underscore the influential role of median filters on sleep
metrics.

\hypertarget{tbl-summary_zm}{}
\begin{longtable}[]{@{}llllllll@{}}
\caption{\label{tbl-summary_zm}Summary of sleep quality metrics as
calculated based on the raw, the 5-minute, and 10-minute median filtered
raw ZM prediction.}\tabularnewline
\toprule\noalign{}
& & Raw & & Median 5-Minute & & Median 10-Minute & \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& & Raw & & Median 5-Minute & & Median 10-Minute & \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
& & Mean & SD & Mean & SD & Mean & SD \\
& SPT & 9.2 & 2.1 & 9.2 & 2.1 & 9.2 & 2.1 \\
& TST & 7.7 & 1.9 & 8.1 & 2.0 & 8.2 & 2.1 \\
& SE & 82.6 & 12.0 & 86.4 & 12.7 & 87.5 & 12.9 \\
& LPS & 34.5 & 27.9 & 36.3 & 39.8 & 38.0 & 48.7 \\
& WASO & 39.0 & 33.6 & 30.6 & 46.8 & 22.3 & 55.4 \\
\end{longtable}

>>>>

\begin{keyword}
    Sleep \sep Accelerometry \sep EEG \sep Machine learning \sep 
    Sleep quality metrics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced]}{\end{tcolorbox}}\fi

@@

In this study, raw data processing began with a low-pass filtration step
using a 4th order Butterworth filter with a 5 Hz cut-off frequency to
eliminate high-frequency noise. Following filtration, data were
partitioned into overlapping 2-second intervals, each successive
interval sharing a 50\% overlap with the previous one similar to methods
described by Skotte et al.\citep{skotte_detection_2014}. Any non-wear
data was remove using previously described methods\citep{skovgaard2023}
and data was resampled to 30-second epochs so every sample classified by
the algorithms corresponds to a 30-second epoch scored during the ZM
recordings. Subsequently, we performed a feature extraction process that
yielded a set of 88 features, providing a robust characterization of the
data. Extracted from accelerometer and temperature signals, these
features include temporal elements that use both lag and lead values,
capturing dynamic data trends by incorporating measurements from
preceding and upcoming intervals. Furthermore, inspired by Walch et
al.\citep{walch2019}, we developed sensor-independent features to
encapsulate circadian rhythms. These features offer unique insights not
directly discernible from sensor outputs and are meant to approximate
the changing drive of the circadian clock to sleep over the course of
the night (see Figure~\ref{fig-sensor-independent}). Furthermore, the
feature set was enriched by including signal characteristics, which
encompass vector magnitude, mean crossing rate, skewness, and kurtosis
for each of the x, y, and z dimensions. These features provide a
comprehensive understanding of the signal's attributes, facilitating a
more nuanced analysis of the underlying patterns. All features are
summarized in table ??? (overvej tabel til supp. material)

@@

\hypertarget{model-validation-evt.-bare-statistics}{%
\subsubsection{Model Validation {[}evt. bare
``statistics''{]}}\label{model-validation-evt.-bare-statistics}}

In our study, we utilized standard evaluation metrics to assess the
performance of each model on an epoch-to-epoch basis. These include
accuracy (\(accuracy = \frac{TP+TN}{TP+TN+FP+FN}\)), sensitivity
(\(sensitivity = \frac{TP}{TP+FN}\)), specificity
(\(specificity = \frac{TN}{TN+FP}\)), precision
(\(precision = \frac{TP}{TP+FP}\)), negative predictive value (NPV,
\(NPV = \frac{TN}{TN + FN}\)), and F1 score
(\(F_1 = 2 * \frac{precision * sensitivity}{precision + sensitivity}\)).

In the context of our sequential learning strategy, the initial models
were tasked with the binary classification of in-bed vs.~out-of-bed. For
this task, we assessed performance using the F1-score, accuracy,
sensitivity, specificity, and precision metrics. The second models in
our sequential learning strategy focused on the binary classification of
asleep vs.~awake. For these models, we considered the same metrics, in
addition to the negative predictive rate. The class imbalance in this
case led us to compute the F1 score as an unweighted macro-average.
Additionally, we evaluated the multiclass classifier, biLSTM, using
macro-averaged F1-score, sensitivity, and precision. Furthermore, we
present precision-recall curves and receiver operating characteristic
curves, including the area under the curve, for each model and for each
type of ZM prediction.We considered both the in-bed/out-of-bed and
awake/asleep scoring tasks as binary classification problems,
designating in-bed and asleep as the positive labels and out-of-bed and
awake as the negative labels in accordance with previous
research\citep{hjorth2012, kushida2001}.

@@

The children children included in the analyses were 9.4 years old (SD =
2.1). The number of sleep epochs recorded in the study was 2,035,261,
which accounted for approximately 86\% of the total sleep duration. This
metric indicates the granularity of sleep measurement, capturing the
discrete moments of sleep throughout the sleep period. Sleep epochs are
essential for analyzing sleep patterns and stages, as well as for
assessing sleep quality and efficiency.

As reported in Table~\ref{tbl-zm_overview} the sleep quality metrics
derived from ZM predictions were modified by the implementation of
5-minute and 10-minute median filters. The Sleep Period Time (SPT) were
consistent across raw and filtered datasets (mean: 9.2 ± 2.1 hours) as
this corresponds to the length of the ZM recording. Total Sleep Time
(TST) and Sleep Efficiency (SE) increased in the filtered data, implying
the filters categorize some wakefulness as sleep. Specifically, TST
increased from a raw mean of 7.7 ± 1.9 hours to 8.1 ± 2.0 hours
(5-minute filter) and 8.2 ± 2.1 hours (10-minute filter), while SE rose
from 82.6 ± 12.0\% to 86.4 ± 12.7\% and 87.5 ± 12.9\% respectively.
Latency to Persistent Sleep (LPS) also elevated, suggesting the filter
smooths out brief awakenings at sleep onset, leading to a prolonged time
to persistent sleep. The most significant change was seen in Wake After
Sleep Onset (WASO), which dropped from 39.0 ± 33.6 minutes in raw data
to 30.6 ± 46.8 minutes and 22.3 ± 55.4 minutes in the 5-minute and
10-minute filtered data, respectively. The results underscore the role
of median filters on sleep metrics.

\hypertarget{tbl-zm_overview}{}
\begin{longtable}{lrrrrrr}
\caption{\label{tbl-zm_overview}Overview of characteristics of the ZM sleep quality summaries. Values
are represented as mean (SD). }\tabularnewline

\toprule
 & SPT (hrs) & TST (hrs) & SE (\%) & LPS (min) & WASO (min) & Awakenings (N) \\ 
\midrule
Raw ZM Predictions & 9.21 (2.15) & 7.74 (1.93) & 82.61 (11.97) & 34.53 (27.87) & 39 (33.56) & 34.46 (11.33) \\ 
5-Min Median & 9.21 (2.15) & 8.1 (2.03) & 86.35 (12.72) & 36.26 (39.8) & 30.63 (46.84) & 4.43 (3.26) \\ 
10-Min Median & 9.21 (2.15) & 8.21 (2.06) & 87.54 (12.89) & 37.99 (48.7) & 22.27 (55.38) & 1.95 (2.01) \\ 
\bottomrule
\end{longtable}

\hypertarget{performance-on-epoch-to-epoch-basis}{%
\subsection{Performance on Epoch-To-Epoch
Basis}\label{performance-on-epoch-to-epoch-basis}}

The epoch-to-epoch evaluation of in-bed performance metrics, outlined in
Table~\ref{tbl-in_bed_performance}, demonstrates practically equivalent
performance across all model types. While the Decision Tree model posted
an F1 score of 94.4\% and accuracy of 95.3\%, the Logistic Regression
and Feed-Forward Neural Network models each exhibited an F1 score of
95.0\% and similar accuracies of 95.7\% and 95.8\%, respectively. The
XGBoost model, despite recording the highest metrics with an F1 score of
95.4\% and accuracy of 96.1\%, outpaced the others only marginally. This
underscores the consistency of performance among these models in
classifying in-bed conditions.

\hypertarget{tbl-in_bed_performance}{}
\begin{longtable}{lrrrrr}
\caption{\label{tbl-in_bed_performance}In-Bed Performance Metrics }\tabularnewline

\toprule
model & F1 Score (\%) & Accuracy (\%) & Sensitivity (\%) & Precision (\%) & Specificity (\%) \\ 
\midrule
Decision Tree & $94.36$ & $95.27$ & $93.12$ & $95.64$ & $96.86$ \\ 
Logistic Regression & $94.99$ & $95.74$ & $95.03$ & $94.94$ & $96.26$ \\ 
Feed-Forward Neural Net & $95.03$ & $95.77$ & $95.07$ & $94.99$ & $96.29$ \\ 
XGBoost & $95.38$ & $96.06$ & $95.83$ & $94.94$ & $96.23$ \\ 
\bottomrule
\end{longtable}

Table~\ref{tbl-sleep_performance} details the performance of the
sequential models on raw and median-filtered (5 and 10 minute) ZM
predictions. The F1 Scores, which are unweighted macro averages, for raw
ZM predictions range from 71.05\% to 76.18\%. The models perform
comparably, but the low Specificity values (62.84\% to 70.93\%) suggest
difficulty in correctly classifying negative instances. Applying
5-minute median filtering improves the performance metrics. The XGBoost
model tops the charts with an F1 Score of 79.22\% and NPV of 74.00\%.
However, Specificity still remains low, with values between 54.68\% and
74.84\% across all models. With 10-minute median filtering, the metrics
improve further. The XGBoost model still leads with an F1 Score of
80.87\% and an NPV of 75.76\%. But, Specificity remains a concern,
ranging from 57.47\% to 76.35\% across all models.

\hypertarget{tbl-sleep_performance}{}
\begin{longtable}{cccccc}
\caption{\label{tbl-sleep_performance}Sleep Performance Metrics }\tabularnewline

\toprule
 & F1 Score (\%) & Precision (\%) & NPV (\%) & Sensitivity (\%) & Specificity (\%) \\ 
\midrule
\multicolumn{6}{l}{Raw ZM Predictions} \\ 
\midrule
Decision Tree & $72.94$ & $93.24$ & $48.36$ & $86.34$ & $67.15$ \\ 
Logistic Regression & $71.05$ & $93.72$ & $43.88$ & $82.72$ & $70.93$ \\ 
Neural Network & $71.76$ & $93.77$ & $45.13$ & $83.59$ & $70.83$ \\ 
XGBoost & $76.18$ & $92.80$ & $57.98$ & $91.32$ & $62.84$ \\ 
\midrule
\multicolumn{6}{l}{5-Min Median} \\ 
\midrule
Decision Tree & $75.48$ & $94.20$ & $55.46$ & $93.35$ & $59.01$ \\ 
Logistic Regression & $68.32$ & $95.84$ & $36.03$ & $81.36$ & $74.84$ \\ 
Neural Network & $71.74$ & $95.78$ & $41.64$ & $85.62$ & $73.11$ \\ 
XGBoost & $79.22$ & $93.87$ & $74.00$ & $97.30$ & $54.68$ \\ 
\midrule
\multicolumn{6}{l}{10-Min Median} \\ 
\midrule
Decision Tree & $76.28$ & $94.74$ & $58.06$ & $94.86$ & $57.47$ \\ 
Logistic Regression & $67.96$ & $96.54$ & $34.29$ & $81.87$ & $76.35$ \\ 
Neural Network & $70.95$ & $96.06$ & $39.54$ & $86.48$ & $71.37$ \\ 
XGBoost & $80.87$ & $94.90$ & $75.76$ & $97.72$ & $57.60$ \\ 
\bottomrule
\end{longtable}

The analysis of precision-recall and ROC curves across different models
and ZM prediction types shows varying performance. In terms of
precision-recall AUC, the Decision Tree model consistently outperforms
others, indicating its superior predictive accuracy (see
Figure~\ref{fig-pr_curves}). Conversely, the Neural Network model
generally shows weaker performance. However, for ROC AUC, the XGBoost
model consistently excels across all data types, indicating a strong
ability to differentiate between classes, while the Neural Network model
tends to underperform (see Figure~\ref{fig-roc_curves}). The F-measure
(F1 score) shows variable performance across different configurations
but generally, the Decision Tree model yields higher scores.

{\centering \includegraphics{visuals/plot_sleep_pr.pdf}

\caption{\label{fig-pr_curves}Precision-recall curves of the models
evaluated across the different ZM predictions, including raw ZM
predictions, as well as 5-minute and 10-minute median smoothing of the
ZM raw predictions. The x-axis of the plot represents the proportion of
true wake epochs that were correctly classified as wake, while the
y-axis represents the proportion of all epochs labeled as wake by the
classifier that were classified correctly. The area under the curve
values are displayed as color-coded text in the plot indicate the area
under the Precision-Recall curve for each model and condition.}

{\centering \includegraphics{visuals/plot_sleep_roc.pdf}

\caption{\label{fig-roc_curves}Receiver operating characteristic curves
of the models evaluated across the different ZM predictions, including
raw ZM predictions, as well as 5-minute and 10-minute median smoothing
of the ZM raw predictions. The x-axis of the plot represents the
proportion of true asleep epochs that were incorrectly classified as
awake, while the y-axis represents the proportion of all epochs labeled
as awake by the classifier that were correctly classified. The area
under the curve values displayed are displayed as color-coded text in
the plot to indicate the area under the receiver operating
characteristic curve for each model and condition.}

The performance metrics of the biLSTM were\ldots{}
